---
title: "Machine Learning - Mid-Term Practice"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(caret)
library(recipes)
library(vip)
```


### ----------------------------------------------------------------------------

## Question 1 

This problem concerns the **College.rds** dataset which contains statistics for a large number of US Colleges from the 1995 issue of US News and World Report. Load the dataset into your R session using the following code.

```{r}
College <- readRDS("College.rds")
```

The dataset contains information on the following variables.

* `Private` - A factor with levels 'No' and 'Yes' indicating private or public university
* `Apps` - Number of applications received
* `Top10perc` - Pct. new students from top 10% of H.S. class
* `Top25perc` - Pct. new students from top 25% of H.S. class
* `F.Undergrad` - Number of full-time undergraduates
* `P.Undergrad` - Number of part-time undergraduates
* `Outstate` - Out-of-state tuition
* `Room.Board` - Room and board costs
* `Books` - Estimated book costs
* `Personal` - Estimated personal spending
* `PhD` - Pct. of faculty with Ph.D.'s
* `Terminal` - Pct. of faculty with terminal degree
* `S.F.Ratio` - Student/faculty ratio
* `perc.alumni` - Pct. alumni who donate
* `Expend` - Instructional expenditure per student
* `Grad.Rate` - Graduation rate
* `Yield` - Percentage of students who choose to enroll in a particular college or university after having been offered admission

Consider `Yield` as the response variable and the rest of the variables as predictors. 

Compare the performance (in terms of **RMSE**) of the following two models:

* A linear regression model;

* A $K$-NN regression model with optimal $K$ chosen by CV. Consider the grid of possible $K$ values given below.


**Perform the following tasks.**

* Investigate the dataset and complete any necessary tasks.

```{r}
# write your R code here



```


* Split the data into training and test sets (80-20).

```{r}
# write your R code here
set.seed(333)



```


* Perform required data preprocessing and create the blueprint. If using `step_dummy()`, set `one_hot = FALSE`. Prepare the blueprint on the training data. Obtain the modified training and test datasets. **Explain the blueprint steps that you use.**

```{r}
# write your R code here
set.seed(333)



```


* Implement 5-fold CV repeated 5 times for each of the models above. 

```{r}
# write your R code here
set.seed(333)




k_grid <- expand.grid(k = seq(1, 101, by = 10))   # for KNN



```


* Report the optimal CV RMSE of each model. Report the optimal value of $K$ for the $K$-NN model. Which model performs better in this situation? 

```{r}
# write your R code here


```

```{}
# write your answer here



```


* Using the optimal model, obtain predictions on the test set. Calculate and report the test set RMSE.

```{r}
# write your R code here



```


* Using the CV object for the linear regression model, obtain the variable importance measures for each feature.

```{r}
# write your R code here



```


### ----------------------------------------------------------------------------


## Question 2

For this question, you will work with the **banknote.rds** dataset. Load the dataset using the following code.

```{r,message=FALSE}
banknote <- readRDS("banknote.rds")
```

The task is to predict whether a banknote is `authentic` or not ("Yes"/"No" variable) based on a number of measures taken from banknote images: `variance`, `skewness`, `kurtosis`, and `entropy` of wavelet transformed images.

Compare the performance (in terms of **Accuracy**) of the following two models:

* Logistic regression;

* $K$-NN classifier with optimal $K$ chosen by CV. Consider the grid of possible $K$ values given below.


**Perform the following tasks.**

* Investigate the dataset and complete any necessary tasks.

```{r}
# write your R code here



```


* Split the data into training and test sets (80-20).

```{r}
# write your R code here
set.seed(333)



```


* Perform required data preprocessing and create the blueprint. If using `step_dummy()`, set `one_hot = FALSE`. Prepare the blueprint on the training data. Obtain the modified training and test datasets. **Explain the blueprint steps that you use.**

```{r}
# write your R code here
set.seed(333)



```


* Implement 5-fold CV repeated 5 times for each of the models above. Use `metric = "Accuracy"`.

```{r}
# write your R code here
set.seed(333)


# CV specifications (To use the code below, remove # and fill the blanks appropriately)
# cv_specs <- trainControl(method = __________, 
#                          number = __________, 
#                          repeats = __________)  




k_grid <- expand.grid(k = seq(1, 101, by = 10))  # for KNN 



```


* Report the optimal CV Accuracy for each model. Report the optimal value of $K$ for the $K$-NN classifier. Which model performs better in this situation? 

```{r}
# write your R code here


```

```{}
# write your answer here



```


* Finally, with the optimal model, obtain class label predictions on the test set (use threshold of 0.5). Create the corresponding confusion matrix and report the test set accuracy.

```{r}
# write your R code here



```


### ----------------------------------------------------------------------------

